Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-07 ltx/paper.tex
pretty much final first draft

\documentclass[smallcondensed, draft]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

@@

\subsubsection{Transition Kernels and Motion Behavior}

@@

\section{Inference}
\label{sec:inference}
Algorithms that infer latent model parameters in a Bayesian manner given the data gained during extraction provide detection and tracking results. By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated and applied independently to gain detection and tracking results. This section provides background on the two inference procedures implemented in this study, both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}. The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in a sequential manner that operates by making use of the time-dependent nature of the model.

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as

@@

where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as

@@

where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}. This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.

@@

% MCMC inference : sampling assignment variables
% ----------------------------------------------

@@

A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments. The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster. The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
\begin{equation}
\begin{split}
p(c_{i,t} = k | \text{state}) & \propto \prod_{t' = t}^{d_{i,t}} P(c_{i,t'} | m_{k,t'-1}, \bold{c}_{1:t'-1}) \\
 & \times
\begin{cases}
	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{split}
\end{equation}
where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
\begin{equation}
P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
\begin{cases}
m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{equation}
which is taken under the situation where $c = k$. Note that the above integral can be determined analytically to give
\begin{equation}
\label{marginal_obs_prob}
\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}()
\end{equation}
where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \willie{cite}.

If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps. \willie{finish writing the few tricky technical stuff here}

% MCMC inference : sampling cluster parameters
% --------------------------------------------

\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
The conjugacy of the chosen distributions in the appearance model and transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
% \begin{eqnarray}
% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
% \end{eqnarray}
\begin{align}
\begin{split}
P(\theta_{k,t} | state) & = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
 & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
\end{split}
\end{align}
Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations in the following Bayesian updates
\begin{eqnarray}
\label{bayesian_update_1}
\kappa_{N} &=& \kappa_{0} + N \\
\label{bayesian_update_2}
\nu_{N} &=& \nu_{0} + N \\
\label{bayesian_update_3}
\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
\label{bayesian_update_4}
\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
\end{eqnarray}
where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
% \begin{eqnarray}
% \overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
% S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
% \end{eqnarray}

% MCMC inference : sampling auxiliary variables
% ---------------------------------------------

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
\begin{equation}
P(z_{k,t,m} | state) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
\end{equation}
This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique. In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample. The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}

@@

% MCMC inference : sampling deletion variables
% --------------------------------------------

\subsubsection{Sampling Deletion Variables $d_{i}$}
Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on (in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}) may be computationally expensive due to the large number of potential deletion times. To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$. As the `lifetime' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
\begin{align}
\begin{split}
l_{i}  &\sim  \text{Geo}(\rho)  \\
d_{i}^{*}  &= l_{i} + t + 1
\end{split}
\end{align}
and accept or reject this sample using the standard M-H acceptance ratio \willie{write computation of ratio}.

% MCMC inference : initializing the sampler
% -----------------------------------------

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of sample of the assignment for each observation $\bold{x}_{1,t}, \ldots, \bold{x}_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept. \willie{Gibbs sampling is used to generate the samples associated with each particle at each time step.}

@@

\begin{algorithm}[!]
\caption{Sequential Monte Carlo Inference for the GPUDDPM}
\label{alg:SMC}
\begin{algorithmic}[1]
\FOR{$l = 1 : L$}
\STATE $w_{0}^{(l)} \leftarrow 1/L$ \hfill $\triangleright$ initialize weights
\ENDFOR
\STATE $K_{0}^{(l)} \leftarrow 0$
\FOR{$t = 1 : T$}
\FOR{$l = 1 : L$}
\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
\STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
\FOR{$s = 1 : S$}
\FOR{$i = 1 : N_{t}$}
\IF{$s=1$}
\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ELSE
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
\ENDIF
\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1$}
\STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\STATE $K_{t}^{(l)} \leftarrow K_{t}^{(l)} + 1$
\STATE $m_{K_{t}^{(l)},t}^{(l)} \leftarrow 1$
\ENDIF
\ENDFOR
\FOR{$k = 1 : K_{t}^{(l)}$}
\IF{$k > K_{t-1}^{(l)}$}
\STATE Sample $\theta_{k,t}^{(l)} \sim q_{1}(\{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
\ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
\STATE Sample $\theta_{k,t}^{(l)} \sim q_{2}(\theta_{k,t-1}^{(l)}, \{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q2_eqn}
\ELSIF{$m_{k,t}^{(l)} > 0$}
\STATE Sample $\theta_{k,t}^{(l)} \sim P(\theta_{k,t}^{(l)} | \theta_{k,t-1}^{(l)})$  \hfill $\triangleright$ eqs.~\eqref{transker1} $\&$ \eqref{transker2}
\ENDIF
\IF{$s=S$}
\STATE Sample $m_{k,t+1}^{(l)} \sim \text{DEL}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
\ENDIF
\ENDFOR
\ENDFOR
\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
\ENDFOR
\FOR {l = 1 : L}
\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$
\ENDFOR
\STATE Resample particles $1, \ldots, L$ and weights $w^{(1)}, \ldots, w^{(L)}$
\ENDFOR
\end{algorithmic}
\end{algorithm}

>>>>

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\label{gpuddpm_def}

@@

\subsubsection{Transition Kernel and Motion Model}

@@

\section{Inference}
\label{sec:inference}
Algorithms that infer latent model parameters via Bayesian inference given the data gained during extraction provide detection and tracking results. By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking. This section provides background on the two inference algorithms implemented in this study, which are both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}. The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from the posterior distribution of the model. The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from the posterior distribution of the model, but in a sequential manner that makes use of the time-dependent nature of the data.

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}. To allow for modeling time-dependent data, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the values of the cluster sizes over time. Instead of incorporating the cluster size variable $m$ directly (as it was in the definition of the GPUDDPM given in \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time at which observations are lost from a cluster. Specifically, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. With these we can express the size of cluster $k$ at time $t$, $m_{k,t}$, as

@@

where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as

@@

where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}. This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.

@@

% mcmc: sampling assignment variables
% -----------------------------------

@@

A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments. The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster. The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
\begin{equation}
\begin{split}
p(c_{i,t} = k | \text{State}) & \propto
\prod_{i'=1}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
 & \times
\begin{cases}
	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{split}
\end{equation}
% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
%\begin{equation}
%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
%\begin{cases}
%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
%\end{cases}
%\end{equation}
where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$. Note that the above integral can be determined analytically to give
\begin{equation}
\label{marginal_obs_prob}
\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
\end{equation}
where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.

%%%%%
% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation
%%%%%

If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables for this new cluster must be initialized for all time steps before sampling can proceed. In experiments, this initialization was carried out by iteratively samping forward to time T and backwards to time 1 via the transition kernel.

% mcmc: sampling cluster parameters
% ---------------------------------

\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
The conjugacy of the chosen distributions in the appearance model and transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
% \begin{eqnarray}
% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
% \end{eqnarray}
\begin{align}
\begin{split}
P(\theta_{k,t} | \text{State}) & = P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
 & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
\end{split}
\end{align}
Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations for the following Bayesian updates
\begin{eqnarray}
\label{bayesian_update_1}
\kappa_{N} &=& \kappa_{0} + N \\
\label{bayesian_update_2}
\nu_{N} &=& \nu_{0} + N \\
\label{bayesian_update_3}
\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
\label{bayesian_update_4}
\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
\end{eqnarray}
where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
% \begin{eqnarray}
% \overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
% S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
% \end{eqnarray}

% mcmc: sampling auxiliary variables
% ----------------------------------

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
\begin{equation}
P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
\end{equation}
This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique. In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample. The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}

@@

% mcmc: sampling deletion variables
% ---------------------------------

\subsubsection{Sampling Deletion Variables $d_{i}$}
Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on (in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}) may be computationally expensive due to the large number of potential deletion times. To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$. As the `lifetime' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
\begin{align}
\begin{split}
l_{i}  &\sim  \text{Geo}(\rho)  \\
d_{i}^{*}  &= l_{i} + t + 1
\end{split}
\end{align}
and accept or reject this sample using the standard M-H acceptance ratio.

%%%%%
% need to include the "standard M-H acceptance ratio"
%%%%%

% mcmc: initializing the sampler
% ------------------------------

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process. Furthermore, within each time step, Gibbs sampling is used to generate the samples associated with each particle.

@@

%%%%%
% in SMC algorithm: still need to include computing r and make sure things are correct
%%%%%
